Diversification of Australian blind snakes

### 2022-11-18 manuscript.Rmd
update talk about biome transitions

## Biogeographic history

We identified the DEC+_j_+_x_ model as the most preferred biogeographical history model (delta AICc = -4.47, see Table-supp-BioGeoBEARS). 
Likelihood ratio test (LRT) indicates that both founder events (+*j* parameter) and distance between biome (+x parameter) together provides statistically significant improvement over the simpler DEC model, DEC+x  and DEC+j (X^2^, DF = 2, D-statistic = 26.7, p-value < 0.001). 
We inferred tropical grassland biome to be the ancestral biome of Australian _Anilios_ (\@ref(fig:fig-biogeobears), but the origin of crown *Anilios* is ambiguous.

## Biome transitions and current distributions

Following the hypothesized colonisation event from a single common ancestor, the diversifcation of _Anilios_ occurred very rapidly. Descendents of the earliest split remained in the Tropical grassland biome but within a million years later another split occured. The earliest biome transition was from tropical grassland into Temperate Forest around 25 MYA (95% HPD: 28 – 22 MYA) and the Arid zone 24 MYA (95% HPD: 28 – 21 MYA). Subsequent splits show transitions from the Arid zone into the Mediterranean biome starting after 15 MYA (\@ref(fig:fig-biogeobears). Current distributions of _Anilios_ are mostly concentrated in the tropical grassland biome closely followed by the Arid zone and Mediterranean biome. We observed that closely related species tend to share the same biome and occur around the same geographical areas.

@@

>>>>

## Biogeographic history and biome transitions

We identified the DEC+_j_+_x_ model as the most preferred biogeographical history model (delta AICc = -4.47, see Table-supp-BioGeoBEARS). 
Likelihood ratio test (LRT) indicates that both founder events (+*j* parameter) and distance between biome (+x parameter) together provides statistically significant improvement over the simpler DEC model, DEC+x  and DEC+j (X^2^, DF = 2, D-statistic = 26.7, p-value < 0.001). 
Our best model inferred tropical grassland biome to be the ancestral biome of Australian _Anilios_ (\@ref(fig:fig-biogeobears)), but the origin of crown *Anilios* is ambiguous.

Following the hypothesized colonisation event from a single common ancestor, the diversifcation of _Anilios_ occurred very rapidly. Descendents of the earliest split remained in the tropical grassland biome but within a million years later another split occured. The earliest biome transition was from tropical grassland biome into the temperate forest biome around 25 MYA (95% HPD: 28 – 22 MYA) for branch leading _A. siliva_, and another branch that included current arid and mediterranean species starting 24 MYA (95% HPD: 28 – 21 MYA). Based on 50 biogeographic stochastic mapping, we found that the largest source of total outward dispersal (anagenetic and cladogenetic) is the tropical grassland biome (642 dispersals, mean = 71.3 per BSM) almost double the number of dispersals from the arid biome (395 dispersals, mean = 43.9 per BSM) (\@ref(fig:fig-biogeobears)).

The top recipients of dispersals from tropical grassland biome is the temperate grassland (5.06 ± 0.79), follotropical forest (4.12 ± 1.29), ,

Subsequent splits show transitions from the arid biome into the mediterranean biome starting after 15 MYA (\@ref(fig:fig-biogeobears). Current distributions of _Anilios_ are mostly concentrated in the tropical grassland biome closely followed by the Arid zone and Mediterranean biome. We observed that closely related species tend to share the same biome and occur around the same geographical areas.

@@

- Limitations is we can only coarsely assign species to one biome one some are range-limited endemics on islands (e.g. _A. longissimus_)
